---
title: "ADNIMERGE2-Analysis-Data"
output: 
   rmarkdown::html_vignette:
     toc: true
     code_folding: hide
date: "Last Updated: `r format(Sys.Date(), '%B %d, %Y')`"
vignette: >
  %\VignetteIndexEntry{ADNIMERGE2-Analysis-Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  class.source = "fold-show"
)

# Data print function
datatable <- function(data, paging = FALSE, searchable = TRUE, bInfo = FALSE, ...) {
  DT::datatable(
    data = data, ...,
    options = list(
      paging = paging, searchable = searchable, bInfo = bInfo,
      ...
    )
  )
}
```

```{r convert-metacore-dataframe, echo = FALSE, class.source = NULL}
# Function to convert `R6-wrapper` metacore objects into a data.frame
convert_metacore_to_dataframe <- function(metacore_data, dataset_name = NULL) {
  require(metacore)
  if (is_metacore(metacore_data) == FALSE) {
    stop("`metacore_data` must be a metacore R6-Wrapper object.")
  }
  dataset_list <- metacore_data$ds_spec$dataset

  metadata <- lapply(dataset_list, function(i) {
    return(metacore_data %>%
      select_dataset(dataset = i, simplify = TRUE))
  }) %>%
    bind_rows() %>%
    left_join(
      metacore_data$ds_spec %>%
        select(dataset, structure, dataset.label = label),
      by = "dataset"
    ) %>%
    {
      if (!is.null(dataset_name)) {
        filter(., dataset %in% dataset_name)
      } else {
        (.)
      }
    }

  return(metadata)
}
```

```{r single-dataset-build-derived, echo = FALSE, class.source = NULL}
# Derived data from a single dataset
single_build_from_derived <- function(metacore, ds_list, dataset_name = NULL, predecessor_only = TRUE, keep = FALSE) {
  require(metacore)
  require(dplyr)
  require(stringr)

  if (length(names(ds_list)) != 1) stop("`ds_list` must be a length of one!")
  input_dataset_name <- names(ds_list)
  input_dataset_name <- toupper(input_dataset_name)
  metacore <- metacore %>%
    select_dataset(.data = ., dataset = dataset_name, simplify = FALSE)

  # Predecessor from a single dataset
  derivations <- metacore$derivations %>%
    filter(str_detect(string = derivation, pattern = paste0("^", input_dataset_name, "\\.")))

  if (nrow(metacore$supp) == 0) {
    supp <- tibble(
      dataset = character(),
      variable = character(),
      idvar = character(),
      qeval = character()
    )
  } else {
    supp <- metacore$supp
  }
  # Internal modified metacores
  internal_metacore <- metacore::metacore(
    ds_spec = metacore$ds_spec,
    supp = supp,
    codelist = metacore$codelist,
    derivations = derivations,
    value_spec = metacore$value_spec,
    var_spec = metacore$var_spec,
    ds_vars = metacore$ds_vars
  )
  data <- build_from_derived(
    metacore = internal_metacore,
    ds_list = ds_list,
    dataset_name = dataset_name,
    predecessor_only = predecessor_only,
    keep = keep
  )
  return(data)
}
```

```{r convert-var-to-fact-wrapper, echo = FALSE, class.source = NULL}
# Wrapper function to apply convert_var_to_fct for multiple variables
convert_var_to_fct_wrapper <- function(data, metacore, var) {
  for (var_name in var) {
    data <- metatools::convert_var_to_fct(
      data = data,
      metacore = metacore,
      var = !!rlang::sym(var_name)
    )
  }
  return(data)
}
```

# Introduction

This article describes creating analysis ready dataset using the [PHARMERVERSE](https://pharmaverse.org/) workflow. It uses some pre-generated standardized dataset and a metadata-specs as input. Please refer to the following articles to see how the derived dataset and metadata-specs are generated. 

* [ADNIMERGE2-Derived-Data](ADNIMERGE2-Derived-Data.html)
* [ADNIMERGE2-Analysis-Meta-Specs](ADNIMERGE2-Analysis-Meta-Specs.html)

In ADNI study, the following selected analysis ready dataset will be created for illustration purpose. 

* Subject-Level Analysis Dataset - ADSL
* Analysis Dataset of Adverse Events - ADAE
* Analysis Dataset of Questionnaire - ADQS

**NOTE** 

* *These anlaysis ready dataset may not be fully complied with tthe CDISC-ADaM standardization. More information about the CDISC-ADaM can be found [here](https://www.cdisc.org/standards/foundational/adam).*

# Load Required R Packages

```{r setup-libraries, class.source = NULL}
library(tidyverse)
library(assertr)
library(metacore)
library(metatools)
library(admiral)
library(admiraldev)
library(xportr)
```

```{r study-library, purl=FALSE}
# ADNI study R data package
library(ADNIMERGE2)
```

# Building Analysis Ready Dataset

```{r prep-meta-data, echo = FALSE}
# Convert `R6-wrapper` metacore object into a data.frame
meta_data_dic <- convert_metacore_to_dataframe(METACORES) %>%
  mutate(
    TYPE = str_to_upper(string = type),
    DERIVED = TRUE,
    FLDNAME = variable
  ) %>%
  select(
    TBLNAME = dataset, CRFNAME = dataset.label, structure,
    FLDNAME, LABEL = label, TYPE, DERIVED, type, variable
  )
```

## Subject-Level Analysis Dataset - ADSL

[`ADSL`](../reference/ADSL.html) dataset contains one records per participant. In ADNI study, [`ADSL`](../reference/ADSL.html) dataset includes only the first records/baseline characteristics of participant when they were enrolled in the study as newly-enrollee or screened to the study for the first time.

```{r adsl-data-dic-print, class.source = NULL, echo = FALSE}
adsl_data_dic <- meta_data_dic %>%
  filter(TBLNAME == "ADSL")

adsl_data_dic %>%
  select(FLDNAME, LABEL, TYPE) %>%
  datatable(data, paging = TRUE)
```

```{r adsl-data-dic, echo = FALSE}
adsl_metacore <- METACORES %>%
  select_dataset(dataset = "ADSL", simplify = FALSE)
```

```{r generate-adsl}
ADSL <- build_from_derived(
  metacore = METACORES,
  dataset_name = "ADSL",
  ds_list = list("DM" = DM),
  predecessor_only = TRUE,
  keep = TRUE
) %>%
  # Create AGE Group - creating categorical group from numeric
  create_cat_var(
    data = .,
    metacore = METACORES,
    ref_var = AGE,
    grp_var = AGEGR1,
    num_grp_var = AGEGR1N
  ) %>%
  # Re-code RACE
  create_var_from_codelist(
    data = .,
    metacore = METACORES,
    input_var = RACE,
    out_var = RACEN,
    decode_to_code = TRUE
  ) %>%
  # Add baseline DX
  derive_vars_merged(
    dataset = .,
    dataset_add = RS,
    filter_add = RSTESTCD == "DX" & RSBLFL == "Yes",
    new_vars = exprs(DX = RSORRES),
    order = NULL,
    mode = NULL,
    by_vars = exprs(STUDYID, USUBJID),
    check_type = "error",
    relationship = "one-to-one"
  ) %>%
  # Convert to factor with Control Terms (CT)
  convert_var_to_fct_wrapper(
    data = .,
    metacore = METACORES %>%
      select_dataset(dataset = "ADSL"),
    var = c("ORIGPROT", "DX", "RACE")
  ) %>%
  # Add baseline amyloid status
  call_derivation(
    dataset = .,
    dataset_add = NV,
    derivation = derive_vars_merged,
    variable_params = list(
      params(
        filter_add = NVTESTCD %in% "AMYSTAT" & NVBLFL == "Yes",
        new_vars = exprs(AMYSTAT = NVSTRESC)
      )
    ),
    by_vars = exprs(USUBJID),
    check_type = "error",
    relationship = "one-to-one"
  ) %>%
  # Add baseline education level and marital status
  call_derivation(
    dataset = .,
    dataset_add = SC,
    derivation = derive_vars_merged,
    variable_params = list(
      params(
        filter_add = SCTESTCD == "PTEDUCAT" & !is.na(SCORRES) & VISITNUM <= 1,
        new_vars = exprs(EDUC = SCSTRESN)
      ),
      params(
        filter_add = SCTESTCD == "PTMARRY" & !is.na(SCORRES) & VISITNUM <= 1,
        new_vars = exprs(MARISTAT = SCORRES)
      )
    ),
    order = exprs(VISITNUM),
    mode = "first",
    by_vars = exprs(STUDYID, USUBJID),
    check_type = "error",
    relationship = "one-to-one"
  ) %>%
  # Add baseline BMI derived from weight and height
  call_derivation(
    dataset = .,
    dataset_add = VS,
    derivation = derive_vars_merged,
    variable_params = list(
      params(
        filter_add = VSTESTCD %in% "WEIGHT" & VSBLFL == "Yes",
        new_vars = exprs(WEIGHT = VSSTRESN)
      ),
      params(
        filter_add = VSTESTCD %in% "HEIGHT" & VSBLFL == "Yes",
        new_vars = exprs(HEIGHT = VSSTRESN)
      )
    ),
    by_vars = exprs(USUBJID),
    check_type = "error",
    relationship = "one-to-one"
  ) %>%
  mutate(
    BMI = round(compute_bmi(height = HEIGHT, weight = WEIGHT), 2),
    BMIBLU = ifelse(!is.na(BMI), "kg/m^2", NA_character_),
    AGEU = ifelse(!is.na(AGEU), "Years", NA_character_)
  ) %>%
  # Create Enrollment Flag
  derive_var_merged_ef_msrc(
    dataset = .,
    by_vars = exprs(STUDYID, USUBJID),
    flag_events = list(
      flag_event(dataset_name = "DM", condition = !is.na(RFSTDTC))
    ),
    source_datasets = list(DM = DM),
    new_var = ENRLFL,
    true_value = "Yes",
    false_value = "No"
  ) %>%
  # Imputation for death date and enrollment end date
  # Required parital date/time, in character format
  rename("EOSDTHC" = EOSDT) %>%
  select(-DTHDT) %>%
  mutate(
    EOSDTHC = as.character(EOSDTHC),
    DTHDTC = as.character(DTHDTC)
  ) %>%
  call_derivation(
    dataset = .,
    derivation = derive_vars_dtm,
    variable_params = list(
      params(dtc = DTHDTC, new_vars_prefix = "DTH"),
      params(dtc = EOSDTHC, new_vars_prefix = "EOS")
    ),
    highest_imputation = "M",
    date_imputation = "mid",
    flag_imputation = "auto",
    min_dates = exprs(ENRLDT)
  ) %>%
  derive_vars_dtm_to_dt(
    dataset = .,
    source_vars = exprs(DTHDTM, EOSDTM)
  ) %>%
  # Remove any columns that are not specified in the meta-specs
  drop_unspec_vars(
    dataset = .,
    metacore = METACORES,
    dataset_name = "ADSL"
  ) %>%
  # Add variable labels
  metatools::set_variable_labels(
    data = .,
    metacore = METACORES,
    dataset_name = "ADSL"
  )
```

```{r checks-adsl}
ADSL <- ADSL %>%
  # Check all variables specified are present and no more
  check_variables(
    data = .,
    metacore = METACORES,
    dataset_name = "ADSL"
  ) %>%
  # Checks all variables with CT only contain values within the CT
  check_ct_data(
    data = .,
    metacore = METACORES %>%
      select_dataset("ADSL"),
    na_acceptable = TRUE,
    omit_vars = c("ENRLFL", "DTHFL")
  ) %>%
  # Orders the columns according to the spec
  order_cols(
    data = .,
    metacore = METACORES,
    dataset_name = "ADSL"
  ) %>%
  # Sort the dataset based on key columns
  sort_by_key(
    data = .,
    metacore = METACORES,
    dataset_name = "ADSL"
  ) %>%
  # Check uniqueness of records by key columns
  check_unique_keys(
    data = .,
    metacore = METACORES,
    datase = "ADSL"
  )
```

```{r checks-adsl-cont}
ADSL <- ADSL %>%
  xportr_metadata(
    .df = .,
    metadata = METACORES %>%
      select_dataset("ADSL"),
    domain = "ADSL",
    verbose = "stop"
  ) %>%
  # Check variable type with specs match
  # xportr_type() %>%
  # Assign variable value length from the meta specs
  xportr_length() %>%
  # Assign variable label from meta specs and
  # checks variable label length for max of 40 characters
  xportr_label() %>%
  # Checks variable format
  xportr_format()
```

## Analysis Dataset of Adverse Events - ADAE

[`ADAE`](../reference/ADAE.html) dataset contains one records per adverse events per participant with the following characteristics.

```{r adae-data-dic-print, class.source = NULL, echo = FALSE}
adae_data_dic <- meta_data_dic %>%
  filter(TBLNAME == "ADAE")

adae_data_dic %>%
  select(FLDNAME, LABEL, TYPE) %>%
  datatable(data, paging = TRUE)
```

```{r adae-data-dic}
adae_metacore <- METACORES %>%
  select_dataset(.data = ., dataset = "ADAE", simplify = FALSE)
```

```{r generate-adae}
# Modified user-defined function
ADAE <- build_from_derived(
  metacore = adae_metacore,
  dataset_name = "ADAE",
  ds_list = list("ADSL" = ADSL, "AE" = AE),
  predecessor_only = TRUE,
  keep = FALSE
) %>%
  # Only participant that had at least one adverse events experience
  filter(USUBJID %in% c(AE$USUBJID)) %>%
  verify(nrow(.) == nrow(AE)) %>%
  # Add adverse events onset and ended date
  derive_vars_merged(
    dataset = .,
    dataset_add = AE %>%
      select(USUBJID, AETERM, AESEQ, AESTDTC, AEENDTC),
    by_vars = exprs(USUBJID, AETERM, AESEQ),
    new_vars = NULL,
    check_type = "error",
    relationship = "one-to-one"
  ) %>%
  # AESTDTC and AEENDTC required to be a character variable
  mutate(
    AESTDTC = as.character(AESTDTC),
    AEENDTC = as.character(AEENDTC)
  ) %>%
  # Add imputed date death date from ADSL
  derive_vars_merged(
    dataset = .,
    dataset_add = ADSL %>%
      select(USUBJID, STUDYID, DTHDT),
    by_vars = exprs(USUBJID, STUDYID),
    new_vars = NULL,
    check_type = "error",
    relationship = "many-to-one"
  ) %>%
  # Derive analysis start and end date
  call_derivation(
    dataset = .,
    derivation = derive_vars_dtm,
    variable_params = list(
      params(
        dtc = AESTDTC, new_vars_prefix = "AST", highest_imputation = "M",
        date_imputation = "first", time_imputation = "first"
      ),
      params(
        dtc = AEENDTC, new_vars_prefix = "AEN", highest_imputation = "M",
        date_imputation = "last", time_imputation = "last",
        max_dates = exprs(DTHDT)
      )
    ),
    flag_imputation = "auto",
    min_dates = exprs(ENRLDT)
  ) %>%
  # Convert into date format
  derive_vars_dtm_to_dt(
    dataset = .,
    source_vars = exprs(ASTDTM, AENDTM)
  ) %>%
  # Derive analysis start/end relative day and
  derive_vars_dy(
    reference_date = ENRLDT,
    source_vars = exprs(ASTDT, AENDT)
  ) %>%
  # Derive analysis duration (value and unit)
  derive_vars_duration(
    new_var = ADURN,
    new_var_unit = ADURU,
    start_date = ASTDT,
    end_date = AENDT,
    in_unit = "days",
    out_unit = "days",
    add_one = TRUE,
    trunc_out = FALSE
  ) %>%
  drop_unspec_vars(
    dataset = .,
    metacore = adae_metacore
  ) %>%
  metatools::set_variable_labels(
    data = .,
    metacore = adae_metacore
  )
```


```{r check-adae}
ADAE <- ADAE %>%
  check_variables(
    data = .,
    metacore = adae_metacore
  ) %>%
  # check_ct_data(
  #   data = .,
  #   metacore = adae_metacore,
  #   na_acceptable = TRUE
  # ) %>%
  order_cols(
    data = .,
    metacore = adae_metacore
  ) %>%
  sort_by_key(
    data = .,
    metacore = adae_metacore
  ) %>%
  check_unique_keys(
    data = .,
    metacore = adae_metacore
  ) %>%
  xportr_metadata(
    .df = .,
    metadata = adae_metacore,
    verbose = "stop"
  ) %>%
  # xportr_type() %>%
  xportr_length() %>%
  xportr_label() %>%
  xportr_format()
```


## Analysis Dataset of Questionnaires - ADQS

[`ADQS`](../reference/ADQS.html) dataset contains one records per questionnaire parameter per visits per participant with the following characteristics. 

```{r adqs-data-dic-print, class.source = NULL, echo = FALSE}
adqs_data_dic <- meta_data_dic %>%
  filter(TBLNAME == "ADQS")

adqs_data_dic %>%
  select(FLDNAME, LABEL, TYPE) %>%
  datatable(data, paging = TRUE)
```

```{r adqs-data-dic}
adqs_metacore <- METACORES %>%
  select_dataset(dataset = "ADQS", simplify = FALSE)
```

```{r adqs-data-prep}
# Assumed all questionnaire parameters are continuous (numeric)
ADQS <- single_build_from_derived(
  metacore = adqs_metacore,
  dataset_name = "ADQS",
  ds_list = list("QS" = QS),
  predecessor_only = TRUE,
  keep = TRUE
) %>%
  # Add variable from ADSL dataset
  derive_vars_merged(
    dataset = .,
    dataset_add = single_build_from_derived(
      metacore = adqs_metacore,
      dataset_name = "ADQS",
      ds_list = list("ADSL" = ADSL),
      predecessor_only = TRUE,
      keep = TRUE
    ),
    by_vars = exprs(USUBJID),
    new_vars = NULL,
    check_type = "error",
    relationship = "many-to-one"
  ) %>%
  verify(nrow(.) == nrow(QS)) %>%
  mutate(ADT = as.Date(ADT)) %>%
  # Add analysis timing variable
  derive_vars_dy(
    reference_date = ENRLDT,
    source_vars = exprs(ADT)
  ) %>%
  mutate(
    AVISIT = if_else(ADT <= ENRLDT, "BASELINE", VISIT),
    AVISITN = if_else(ADT <= ENRLDT, 0, VISITNUM)
  ) %>%
  # Re-code PARAMN
  # Assumed all PARAMCD are coded in the meta-specs
  create_var_from_codelist(
    data = .,
    metacore = adqs_metacore,
    input_var = PARAMCD,
    out_var = PARAMN,
    decode_to_code = TRUE
  ) %>%
  # Flag baseline records if the flag is not presented in QS
  restrict_derivation(
    derivation = derive_var_extreme_flag,
    args = params(
      by_vars = exprs(STUDYID, USUBJID, PARAMCD),
      order = exprs(ADT),
      new_var = ABLFL,
      mode = "last"
    ),
    filter = !is.na(AVAL) & ADT <= ENRLDT
  ) %>%
  # Derive baseline and change from baseline variables
  # Only applicable for continuous parameters
  derive_var_base(
    by_vars = exprs(STUDYID, USUBJID, PARAMCD),
    source_var = AVAL,
    new_var = BASE
  ) %>%
  # Derive change for post-baseline records
  restrict_derivation(
    derivation = derive_var_chg,
    filter = AVISITN > 0
  ) %>%
  # Derive percentage change for post-baseline records
  restrict_derivation(
    derivation = derive_var_pchg,
    filter = AVISITN > 0
  ) %>%
  # Derive sequence number
  derive_var_obs_number(
    by_vars = exprs(STUDYID, USUBJID, PARAMCD),
    order = exprs(PARAMCD, ADT),
    new_var = ASEQ,
    check_type = "none" # "error"
  ) %>%
  drop_unspec_vars(
    dataset = .,
    metacore = adqs_metacore
  ) %>%
  metatools::set_variable_labels(
    data = .,
    metacore = adqs_metacore
  )
```

```{r check-adadas}
ADQS <- ADQS %>%
  check_variables(
    data = .,
    metacore = adqs_metacore
  ) %>%
  check_ct_data(
    data = .,
    metacore = adqs_metacore,
    na_acceptable = TRUE,
    omit_vars = c("ENRLFL", "ABLFL", "PARAMCD")
  ) %>%
  order_cols(
    data = .,
    metacore = adqs_metacore
  ) %>%
  sort_by_key(
    data = .,
    metacore = adqs_metacore
  ) %>%
  # check_unique_keys(
  #   data = .,
  #   metacore = adqs_metacore
  # ) %>%
  xportr_metadata(
    .df = .,
    metadata = adqs_metacore,
    verbose = "stop"
  ) %>%
  # xportr_type() %>%
  xportr_length() %>%
  xportr_label()
```


### Analysis Dataset of ADAS-Cog Behaviour - ADADAS

Suppose the ADAS-Cog behaviors 13-item total score was the primary outcome in the analysis, then ADAS-Cog Behavior Analysis Dataset (`ADADAS`) can be created either directly from [`ADQS`](../reference/ADQS.html) or with a dataset-specific metadata-specs (i.e. [adding here](../reference/METACORES.html)). The following chuck illustrate how the `ADADAS` can be created directly from pre-generated [`ADQS`](../reference/ADQS.html).

```{r generate-adadas}
ADADAS <- ADQS %>%
  filter(PARAMCD == "TOTALMOD")
```
