---
title: "ADNIMERGE2-PACC-SCORE"
output: 
  rmarkdown::html_vignette:
    toc: true
    code_folding: hide
date: "Last Updated: `r format(Sys.Date(), '%B %d, %Y')`"
vignette: >
  %\VignetteIndexEntry{ADNIMERGE2-PACC-SCORE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: "`r system.file('references.bib', package = 'ADNIMERGE2')`"
biblio-style: "apalike"
link-citations: true
nocite: |
  @donohue2014preclinical @donohue2017association
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  eval = TRUE
)
# Data print function
datatable <- function(data, paging = FALSE, searchable = TRUE, bInfo = FALSE, ...) {
  DT::datatable(
    data = data, ...,
    options = list(paging = paging, searchable = searchable, bInfo = bInfo, ...)
  )
}
```

# Introduction

This article describes the procedure to calculate a modified version of the Preclinical Alzheimer Cognitive Composite [PACC](https://doi.org/10.1001/jamaneurol.2014.803) score in ADNI study based on @donohue2014preclinical approach.

According to @donohue2014preclinical, the `PACC` score is a baseline standardize composite of:

* Free and Cued Selective Reminding Test (FCSRT) score

* Logical Memory IIa Delayed Recall (LM) score

* Digit Symbol Substitution Test (DSST) score, and 

* Mini-Mental State Examination (MMSE) total score

Each of component score is required to be standardized by the corresponding baseline score prior to computing the composite score. Higher values of the composite score indicate better performance.

In ADNI study, some of these components are either collected partially during certain study phase or not available at all. Since FSCRT is not available at all, the Delayed Recall portion of the Alzheimer's Disease Assessment Scale (ADAS), which is called `ADAS-Cog Q4 score`, will be used as a proxy. The ADAS-Cog Q4 score ranges from `0 to 10` points where higher value indicates worse condition. Similarly, a logarithm scale of `Trails B` score will be used as a proxy for DSST since DSST was collected in the early phase of the study (i.e. `ADNI1`). Therefore, we will have two modified version of `PACC` composite score: 

a. **mPACCdigit: ** Compute when DSST score is used.

b. **mPACCtrailsB: ** Compute when Trails B score is used.

# Load Required R Packages 

```{r setup-pkgs}
library(tidyverse)
library(assertr)
library(DT)
library(labelled)
```

```{r study-pkg, purl = FALSE}
# ADNI study R data package
library(ADNIMERGE2)
```

# Compute PACC Composite Score

```{r pacc-data-dict, eval = TRUE, include = FALSE, echo = FALSE}
# PACC - data dictionary
pacc_data_dic <- bind_rows(
  c(
    FLDNAME = "ORIGPROT",
    LABEL = "Original Study Protocol",
    TEXT = "First time participation in ADNI study phase"
  ),
  c(
    FLDNAME = "COLPROT",
    LABEL = "Study Data Collection Protocol"
  ),
  c(
    FLDNAME = "RID",
    LABEL = "Subject Identifier"
  ),
  c(
    FLDNAME = "VISCODE",
    LABEL = "Study visit code"
  ),
  c(
    FLDNAME = "ADASQ4SCORE",
    LABEL = "Delayed Word Recall Score - portion of ADAS-Cog"
  ),
  c(
    FLDNAME = "ADASQ4SCORE_VISDATE",
    LABEL = "ADAS-Cog Collection/Visit Date"
  ),
  c(
    FLDNAME = "MMSE",
    LABEL = "Mini-Mental Examination Total Score"
  ),
  c(
    FLDNAME = "MMSE_VISDATE",
    LABEL = "MMSE Collection/Visit Date"
  ),
  c(
    FLDNAME = "LDELTOTL",
    LABEL = "Logical Memory - Delayed RecallScore"
  ),
  c(
    FLDNAME = "LDELTOTL_VISDATE",
    LABEL = "LDELTOTL Collection/Visit Date"
  ),
  c(
    FLDNAME = "DIGITSCR",
    LABEL = "Digit Symbol Substitution Score"
  ),
  c(
    FLDNAME = "DIGITSCR_VISDATE",
    LABEL = "DIGITSCR Collection/Visit Date"
  ),
  c(
    FLDNAME = "TRABSCOR",
    LABEL = "Trial B score"
  ),
  c(
    FLDNAME = "LOG.TRABSCOR",
    LABEL = "Trial B score - Log Scale"
  ),
  c(
    FLDNAME = "TRABSCOR_VISDATE",
    LABEL = "TRABSCOR Collection/Visit Date"
  ),
  c(
    FLDNAME = "PACC_VISDATE",
    LABEL = "PACC Collection/Visit Date",
    TEXT = paste0(
      "Either common date across non-missing PACC component ",
      "collection date, the maximum relative date or the corresponding ",
      "study visit date when all component collection date is not available."
    )
  ),
  c(
    FLDNAME = "ADASQ4SCOREZ",
    LABEL = "ADAS Q4 Standardized Score",
    TEXT = paste0(
      "Standardized by corresponding baseline score among ",
      "cognitive normal (CN) new enrollee in ADNI study. ",
      "Also adjusted for directional effect (i.e. multiplied by -1)"
    )
  ),
  c(
    FLDNAME = "MMSEZ",
    LABEL = "Mini-Mental State Examiniation Standardized Score",
    TEXT = paste0(
      "Standardized by corresponding baseline score among ",
       "cognitive normal (CN) new enrollee in ADNI study."
    )
  ),
  c(
    FLDNAME = "LDELTOTLZ",
    LABEL = "Logical Memory - Delayed Recall Standardized Score",
    TEXT = paste0(
      "Standardized by corresponding baseline score among ",
      "cognitive normal (CN) new enrollee in ADNI study."
    )
  ),
  c(
    FLDNAME = "DIGITSCRZ",
    LABEL = "Digit Symbol Substitution Standardized Score",
    TEXT = paste0(
      "Standardized by corresponding baseline score among ",
       "cognitive normal (CN) new enrollee in ADNI1 study phase."
    )
  ),
  c(
    FLDNAME = "LOG.TRABSCORZ",
    LABEL = "Logical Memory - Delayed Recall Log Standardized Score",
    TEXT = paste0(
      "Standardized by corresponding log transformed baseline score among ",
       "cognitive normal (CN) new enrollee in ADNI study. ",
      "Also adjusted for directional effect (i.e. multiplied by -1)"
    )
  ),
  c(
    FLDNAME = "mPACCdigit",
    LABEL = "Modified PACC - Digit Score"
  ),
  c(
    FLDNAME = "mPACCtrailsB",
    LABEL = "Modified PACC - Trial B Score"
  )
) %>%
  tibble::as_tibble() %>%
  relocate(c(FLDNAME, LABEL, TEXT)) %>%
  mutate(
    TEXT = replace_na(TEXT, " "),
    DERIVED = TRUE,
    TBLNAME = "PACC",
    CRFNAME = "Modified Preclinical Alzheimer’s Cognitive Composite"
  )

names(pacc_data_dic$LABEL) <- pacc_data_dic$FLDNAME
```

## Import Raw Datasets

First, we import pre-processed raw datasets that contain PACC component from the `ADNIMERGE2` package. These raw datasets were stored in `csv` format as external data in the package. The reasons for sourcing those dataset is 1) not all PACC component score are yet available at the data-sharing platform, particularly the ADAS-Cog Q4 score, and 2) to reduced any potential additional package dependency.

In the next r chunk, all raw datasets will be loaded to the current working environment using `load_csv_files()` function. 

```{r import-pacc-raw-input, purl = FALSE}
# Import PACC input raw datasets
pacc_raw_data_dir <- system.file(
  "extradata/pacc-raw-input",
  package = "ADNIMERGE2",
  mustWork = TRUE
)
load_csv_files(input_dir = pacc_raw_data_dir)
```

These raw datasets are in a long format and presented as follow:

* `pacc_mmse_long`: A dataset contains the `MMSE` score downloaded from the data-sharing platform.

* `pacc_neurobat`: A dataset contains Logical Memory - Delayed Recall (`LDELTOTL`), Digit Symbol Substitution (`DIGITSCR`), and Trials B (`TRABSCOR`) raw scores from Neuropsychological Battery (`NEUROBAT`) data. The `NEUROBAT` data is also downloaded from the data-sharing platform. Both a long and wide format of the data are loaded.

* `pacc_adas_q4score_long`: A dataset contains the raw `ADAS-Cog Q4` score that obtained from two sources: 
   
   + Current study phase EDC portal
   
   + Old version of ADNI study R data package `ADNIMERGE`, which is also available at the data-sharing platform, that contain records from ADNI1 to ADNI3 study phases.

## Data Preparation Assumption

* During the raw dataset preparation, the screening fail visit code in `ADNI1` study phase was remapped to the actual screening visit code.

* Since some of the PACC component assessments were either not collected at baseline visit during the whole study or partially collected at baseline visit in certain study phase. It is reasonable to carrying forward the screening visit record of an assessment as a corresponding baseline record if the baseline assessment result is missing/unknown. For instance, the `MMSE` was not collected at baseline visit for the whole study phase, and the other neuropsychological assessments (i.e., `LDELTOTL`, `DIGITSCR`, and `TRABSCOR`) are partially collected at baseline visit in certain study phase. 

The following r chuck demonstrate how the screening-baseline visit adjustment is preformed by carrying forward any screening visit record as a baseline record using `adjust_scbl_record()` function.

```{r adjust-scbl}
# Adjustment of screening-baseline record ----
## MMSE ----
pacc_mmse_long <- pacc_mmse_long %>%
  bind_rows(
    adjust_scbl_record(pacc_mmse_long) %>%
      filter(VISCODE %in% get_baseline_vistcode())
  ) %>%
  set_as_tibble()

## NEUROBAT ----
# Required to keep the same records per visit
neurobat_vars <- c("LDELTOTL", "DIGITSCR", "TRABSCOR")
join_by_vars <- c("COLPROT", "RID", "VISCODE", "SCORE_SOURCE")

# Identify carried baseline records from screening visit
adjusted_bl_neurobat <- pacc_neurobat_wide %>%
  adjust_scbl_record() %>%
  pivot_longer(
    cols = all_of(neurobat_vars),
    names_to = "SCORE_SOURCE",
    values_to = "SCORE"
  ) %>%
  # Only baseline records
  filter(VISCODE %in% get_baseline_vistcode()) %>%
  mutate(ADJUSTED_BL = "Yes")

adjusted_bl_neurobat <- adjusted_bl_neurobat %>%
  left_join(
    pacc_neurobat_long %>%
      filter(VISCODE %in% get_baseline_vistcode()) %>%
      mutate(IS_ORGINAL_BL_RECORD = "Yes") %>%
      select(COLPROT, RID, VISCODE, SCORE_SOURCE, IS_ORGINAL_BL_RECORD),
    by = join_by_vars
  ) %>%
  assert_uniq(all_of(join_by_vars)) %>%
  filter(is.na(IS_ORGINAL_BL_RECORD)) %>%
  select(-IS_ORGINAL_BL_RECORD, -ADJUSTED_BL)

pacc_neurobat_long <- pacc_neurobat_long %>%
  bind_rows(adjusted_bl_neurobat) %>%
  distinct() %>%
  assert_uniq(all_of(join_by_vars)) %>%
  set_as_tibble()

# ADAS-cog Q4 Score ----
pacc_adas_q4score_long <- pacc_adas_q4score_long %>%
  set_as_tibble()
```

The next step is to combine all the raw datasets and converting the combined data into a wide format using the study data collection phase/protocol and study visit code per subject.

```{r bind-pacc-input}
# Bind PACC input raw-data
pacc_input_data_long <- bind_rows(
  pacc_mmse_long,
  pacc_neurobat_long,
  pacc_adas_q4score_long
)

# Convert into wide format
pacc_component <- c("ADASQ4SCORE", "MMSE", neurobat_vars)

pacc_input_data_wide <- pacc_input_data_long %>%
  unite(col = "SCORE_VISDATE", SCORE, VISDATE, sep = "-/") %>%
  pivot_wider(
    names_from = SCORE_SOURCE,
    values_from = SCORE_VISDATE
  ) %>%
  assert_uniq(all_of(c("COLPROT", "RID", "VISCODE"))) %>%
  mutate(across(all_of(pacc_component), ~ strsplit(as.character(.), "-/"))) %>%
  unnest_wider(all_of(pacc_component), names_sep = "_") %>%
  rename_with(~ str_remove_all(.x, "\\_1$"), starts_with(pacc_component)) %>%
  rename_with(~ str_replace_all(.x, "2$", "VISDATE"), starts_with(pacc_component)) %>%
  mutate(across(starts_with(pacc_component), ~ ifelse(.x %in% "NA", NA, as.character(.x)))) %>%
  mutate(across(all_of(pacc_component), as.numeric))
```

Some of the PACC component might not be collected on the same date although those assessments could be collected at the same study visit. In order to have a representative common date associated with the PACC score per study visit, we compare the specific study visit date with corresponding component collection date. We use `get_vars_common_date()` function to get a common date per specific study visit per subject as shown in the following two r chunks. A maximum date will be selected if any of the component specific dates are different. A maximum date selection is used due to the fact that some of the screening records were carried forward as a baseline record.

```{r pacc-visit-date}
# Add visit date
pacc_input_data_wide <- pacc_input_data_wide %>%
  left_join(
    REGISTRY %>%
      select(COLPROT, RID, VISCODE, EXAMDATE) %>%
      set_as_tibble(),
    by = c("COLPROT", "RID", "VISCODE")
  ) %>%
  verify(nrow(.) == nrow(pacc_input_data_wide))
```

```{r pacc-common-date}
# Get common date per record
pacc_input_data_wide <- get_vars_common_date(
  .data = pacc_input_data_wide,
  date_cols = paste0(pacc_component, "_VISDATE"),
  compared_ref_date = TRUE,
  ref_date_col = "EXAMDATE",
  select_method = "max",
  preferred_date_col = "COMMON_DATE"
) %>%
  rename("PACC_VISDATE" = FINAL_DATE)
```

## Component Baseline Score Summaries 

The next step is to calculate component specific baseline summaries that will be used to standardized/normalized each component score. The baseline score summary is related to user-defined standardization population. Here, we define the standardization population as follows:

* New enrolled subject in ADNI study

* Subjects with baseline Cognitive Normal (`CN`) diagnostics status

The following r chunk shows how the PACC component baseline score is summarized across the baseline diagnostics status based on the standardization population defined in the above. The summary result is generated using `get_score_summary_stats()` function.

**NOTE: ** Required to apply a logarithm transformation of `Trials B` score if the score is not already in a logarithm scale. 

```{r pacc-component-bl-summary}
pacc_component_bl <- pacc_input_data_wide %>%
  mutate(TRABSCOR = log(TRABSCOR + 1)) %>%
  # Add enrollment date
  left_join(
    get_adni_enrollment(.registry = REGISTRY) %>%
      mutate(RID = as.character(RID)) %>%
      select(ORIGPROT, RID, ENRLDT = EXAMDATE, ENRLFG),
    by = "RID"
  ) %>%
  # Add baseline diagnostics status
  left_join(
    get_adni_blscreen_dxsum(
      .dxsum = DXSUM,
      phase = "Overall",
      visit_type = "baseline"
    ) %>%
      mutate(RID = as.character(RID)) %>%
      select(RID, DX = DIAGNOSIS, DXDATE = EXAMDATE) %>%
      assert_uniq(RID),
    by = "RID"
  ) %>%
  # Baseline visit for new enrollee
  filter(ENRLFG %in% "Yes") %>%
  group_by(RID) %>%
  filter(ORIGPROT == COLPROT) %>%
  filter(VISCODE %in% get_baseline_vistcode()) %>%
  ungroup() %>%
  assert_uniq(RID)

pacc_component_bl.summary <- get_score_summary_stats(
  .data = pacc_component_bl,
  wideFormat = TRUE,
  scoreVar = pacc_component,
  groupVar = "DX",
  filterGroup = "CN"
)
pacc_component_bl.summary
```

## Compute PACC Score 

The composite PACC score is the sum of the standardized component score adjusted by the number of components. The following two conditions should be checked before computing the PACC score.

* `Trials B` Measurement Scale: To identify whether the score is in a logarithm scale. For instance here, the `Trials B` score is not in a logarithm scale. We need either to map the raw score values to logarithm scale prior to using the scoring function or to specify an argument that allows to perform a logarithm transformation inside the scoring function.

* Standardized Component Directional Effect: To verify the same directional effect across standardized component score and convert into the same effect since the composite score will be the sum of the standardized component scores. In ADNI study, `ADAS-cog Q4` and `Trials B` scores have opposite directional effect compared to the remaining PACC component scores. Higher values of those two components indicates worse condition. It required to convert the direction effect of these standardized components. Thus, higher values of the composite score indicates better performance.

Moreover, the composite score will only be calculated if there are at least two non-missing component scores at a given study visit per subject.

Then, we use `compute_pacc_score()` function to compute the two modified composite PACC scores in ADNI study. 

The next r chunk shows how to calculate the PACC composite score based on a wide format input component score, a standardization baseline summaries, and with a logarithm transformation of `Trials B` score inside in the function.

<!-- **NOTE:** Also required to verify the `Trials B` score measurement scale.  -->

<!-- Here, the `Trials B` score is not in logarithm scale. We need either to map the raw score values to logarithm scale prior to using the scoring function (i.e, `compute_pacc_score`) or to specify an argument that allows to perform a logarithm transformation inside the scoring function. -->

```{r compute-pacc-score}
PACC <- compute_pacc_score(
  .data = pacc_input_data_wide,
  bl.summary = pacc_component_bl.summary,
  componentVars = pacc_component,
  rescale_trialsB = TRUE,
  keepComponents = TRUE,
  wideFormat = TRUE
) %>%
  mutate(RID = as.numeric(RID)) %>%
  rename_with(~ str_replace_all(.x, "\\.zscore", "Z"), ends_with(".zscore")) %>%
  create_orig_protocol()

PACC <- PACC %>%
  select(all_of(pacc_data_dic$FLDNAME)) %>%
  labelled::set_variable_labels(
    .labels = pacc_data_dic$LABEL,
    .strict = TRUE
  )
```

<details>
<summary>Show [`PACC`](../reference/PACC.html) Variables</summary>

```{r pacc-data-dic-print, echo = FALSE}
pacc_data_dic %>%
  select(FLDNAME, LABEL, TEXT) %>%
  datatable(., paging = TRUE)
```

</details>

# References {-}

<div id="refs"></div>
